Toward theories of deliberated preferences

### 2020-12-09 simple.tex
Attempt about falsification

A claim $\gC \subseteq \powerset{I × \allhist × \Phi}$ represents the theory’s claim about the relation $\leadsto$: the theory claims that ${\leadsto} \in \gC$. Such a claim is required to be falsifiable, which is defined as follows.
Observations of the relation $\leadsto$ can in general include positive statements, of the form $\hist \ileadsto \phi$, and negative ones, of the form $\hist \nileadsto \phi$. (One way to obtain such negative observations is to use an incompatible proposition $\phi'$ and deduce from $\hist \ileadsto \phi'$ that $\hist \nileadsto \phi$.) A relation $R \subseteq I × \allhist × \Phi$ is compatible with the positive observations $O \subseteq I × \allhist × \Phi$ and the negative observations $O' \subseteq I × \allhist × \Phi$ iff $O \subseteq R \land R \cap O' = \emptyset$.
The claim $\gC$ is said to be falsifiable iff there exists finite set of observations $O, O' \subseteq I × \allhist × \Phi$ such that $\forall R \subseteq I × \allhist × \Phi \suchthat O \subseteq R \land R \cap O' = \emptyset$: $R \notin \gC$. 
In such a case, no possible relation $R$ compatible with the observations $O, O'$ are in $\gC$; in other words, the observations $O, O'$ prove that the claim that ${\leadsto} \in \gC$ is false.

>>>>

A claim $\gC \subseteq \powerset{I × \allhist × \Phi}$ represents the theory’s claim about the relation $\leadsto$: the theory claims that ${\leadsto} \in \gC$. Such a claim is required to be falsifiable. Defining falsification formally require to consider the relationship between the real relation $\leadsto$ and what one can know about it through observations.

Recall that $\leadsto$ will in general not be known in its entirety, but can be queried through observations. Observations of the relation $\leadsto$ can in general include positive statements, of the form $\hist \ileadsto \phi$, and negative ones, of the form $\hist \nileadsto \phi$. (One way to obtain such negative observations is to use an incompatible proposition $\phi'$ and deduce from $\hist \ileadsto \phi'$ that $\hist \nileadsto \phi$.) 
Assuming (possibly counterfactually) that $R \subseteq I × \allhist × \Phi$ is the “real” relation $\leadsto$, the possible observations are $(O, O') \suchthat O, O' \subseteq I × \allhist × \Phi \land [O \subseteq R] \land [R \cap O' = \emptyset]$; and conversely, once we have gathered some observations $(O, O')$, we know that ${\leadsto} \in \set{R \subseteq I × \allhist × \Phi \suchthat O \subseteq R \land R \cap O' = \emptyset}$.

A claim about ${\leadsto}$ being included in some given set is said to be falsifiable iff either it is true, or it is possible to prove it false using a finite set of observations. Formally, the claim $\gC \subseteq \powerset{I × \allhist × \Phi}$ is falsifiable iff $\forall R \notin \gC: \exists \text{ finite sets } O, O' \subseteq I × \allhist × \Phi \suchthat \forall R' \subseteq I × \allhist × \Phi \suchthat O \subseteq R' \land R' \cap O' = \emptyset: R' \notin \gC$. 
In such a case, no possible relation $R'$ compatible with the observations $O, O'$ are in $\gC$; in other words, the observations $O, O'$ prove that the claim that ${\leadsto} \in \gC$ is false.
